Knitr Assignment
====================  
  
  
  
### Introduction

I will use this document to analyse the Activity Monitoring dataset.


  
### Q1 - Code for reading in the dataset and/or processing the data  
```{r}
data<-read.csv("C:/activity.csv")
```
  

### Q2 - Histogram of the total number of steps taken each day  
```{r,results="asis",message=FALSE}
library(dplyr)
dailytotal <- subset(data,is.na(steps)==FALSE) %>% group_by(date) %>% summarise(Total_Steps = sum(steps))
totals<-as.numeric(dailytotal$Total_Steps)
hist(totals, main="Histogram of Total number of steps taken each day", xlab= "Steps Taken")
```
  
  
### Q3 - Mean and median number of steps taken each day  
```{r,results="asis",message=FALSE,warning = FALSE}
library(dplyr)
dailymean <- subset(data,is.na(steps)==FALSE) %>% group_by(date) %>% summarise(Mean = mean(steps))
dailymedian <- subset(data,is.na(steps)==FALSE) %>% group_by(date) %>% summarise(Median = median(steps))
mean<-mean(dailymean$Mean)
median<-median(dailymedian$Median)
library(kableExtra)
k<-kbl(merge(dailymean,dailymedian, by="date"))
```
The average steps is  `r mean` and the median steps is `r median`.
  
### Q4 - Time series plot of the average number of steps taken
```{r,message=FALSE,warning=FALSE}
plot(data=dailymean,y=dailymean$Mean,x=as.Date(dailymean$date),xlab="Date",
     ylab="Average Steps",main="Average steps taken over time")
```

### Q5 - The 5-minute interval that, on average, contains the maximum number of steps  
```{r,message=FALSE,warning=FALSE}
m<-subset(data,is.na(steps)==FALSE) %>% group_by(interval) %>% summarise(Mean = mean(steps))
m2<-head(m[order(-m$Mean),],1)
m3<-m2$interval
```
The 5-minute interval with the max number of steps, on average, is `r m3`.  


### Q6 - Code to describe and show a strategy for imputing missing data  
The below code is used to work out which rows have missing step counts.  
The average, where data isn't missing, is calculated for those intervals
  with missing step counts.  
The imputed averages are joined to the missing data set and stored in a new variable.
```{r,message=FALSE,warning=FALSE}
missing<-subset(data,is.na(steps)==TRUE)
imputed_data<-merge(missing,m,by="interval")
```

### Q7 - Histogram of the total number of steps taken each day after missing values are imputed
```{r,message=FALSE,warning=FALSE}
imputed<-imputed_data[,c("interval","date","Mean")]
colnames(imputed)<-c("interval","date","steps")
alldata<-union(subset(data,is.na(steps)==FALSE),imputed)
dailyimputedtotal <- alldata %>% group_by(date) %>% summarise(Total_Steps = sum(steps))
imputedtotals<-as.numeric(dailyimputedtotal$Total_Steps)
hist(imputedtotals, main="Histogram of Total number of steps taken each day", xlab= "Steps Taken")
dailymean <- alldata %>% group_by(date) %>% summarise(Mean = mean(steps))
dailymedian <-alldata %>% group_by(date) %>% summarise(Median = median(steps))
mean<-mean(dailymean$Mean)
median<-median(dailymedian$Median)
```
The new average steps is  `r mean` and the median steps is `r median`.
  

### Q8 - Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r,message=FALSE,warning=FALSE}
alldata<-union(subset(data,is.na(steps)==FALSE),imputed)
alldata$day<-weekdays(as.Date(alldata$date))
weekendmean <- subset(alldata,day =="Saturday" | day == "Sunday") %>% group_by(interval) %>% summarise(Mean = mean(steps))
weekdaymean <- subset(alldata,day !="Saturday" & day != "Sunday") %>% group_by(interval) %>% summarise(Mean = mean(steps))
par(mfrow=c(1,2))
min<-min(union(weekdaymean$Mean,weekendmean$Mean))
max<-max(union(weekdaymean$Mean,weekendmean$Mean))
plot(data=weekendmean,x=weekendmean$interval,y=weekendmean$Mean,xlab="Interval",ylab="Average steps taken",main="Weekend Data",ylim=c(min,max))
plot(data=weekdaymean,x=weekdaymean$interval,y=weekdaymean$Mean,xlab="Interval",ylab="Average steps taken",main="Weekday Data",ylim=c(min,max))

```
